---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rjsonschema

<!-- badges: start -->
[![R-CMD-check](https://github.com/DavZim/rjsonschema/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DavZim/rjsonschema/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/rjsonschema)](https://CRAN.R-project.org/package=rjsonschema)
<!-- badges: end -->

The goal of `rjsonschema` is to validate JSON strings or objects against JSON schema definitions.
For an introduction to JSON schemas, visit: <https://json-schema.org/>.

The package is a light wrapper around the [`jsonschema-rs`](https://docs.rs/jsonschema/latest/jsonschema/) ([Github source](https://github.com/Stranger6667/jsonschema)) Rust crate.

## Installation

You can install the development version of `rjsonschema` like so:

``` r
# DEV version
# remotes::install_github("DavZim/rjsonschema")

# CRAN version
install.packages("rjsonschema")
```

## Usage Example

```{r example}
library(rjsonschema)

# 1. define a simple schema
schema <- '{"type": "string"}'

# 2. test the schema against JSON strings
# note '123' is a JSON string, which represents an integer and not a string!
validate(schema = schema, instance = '"hello world"') # passes as "hello" is a string
validate(schema = schema, instance = '123') # fails as 123 is not a string

# multiple instances work!
validate(schema = '{"type": "string"}', instance = c('"hello"', '123'))

# get more information about the validation with style basic
validate(schema = '{"type": "string"}', instance = '123', style = "basic") |> 
  str()

validate(schema = '{"type": "string"}', instance = c('"hello"', '123'), style = "basic") |> 
  str()
```

Note, both instance and schema can be JSON strings (as shown above), file names to JSON files, or R objects (mostly lists).

You can also validate a schema against a specific JSON meta schema (the JSON schema definition of a schema).

```{r validate_schema}
schema <- '{"type": "string"}'
# if you want to have a look at the meta scheme, see the following file:
# system.file("schema-draft202012.json", package = "rjsonschema") or
# <https://json-schema.org/draft/2020-12/schema>
validate_schema(schema, version = "draft202012")
```


If you want to use this function with `stopifnot()`, or similar assertion libraries, you can set `err_on_invalid` to `TRUE`:

```{r stopifnot, error=TRUE}
validate(schema = '{"type": "string"}', instance = '123', err_on_invalid = TRUE)
```


## Full Example

A longer example, taken from [the official getting started guide](https://json-schema.org/learn/getting-started-step-by-step), is the following.
For a store, we are given the following schema, which defines a JSON for a product:

```{r, echo=FALSE,results="asis"}
cat("```json\n")
cat(paste(readLines(system.file("example_schema_catalog.json", package = "rjsonschema")), collapse = "\n"))
cat("\n```\n")
```

The schema is shipped with the package and can be accessed like so:

```{r schema_load}
schema <- system.file("example_schema_catalog.json", package = "rjsonschema")
```

We can now validate a JSON object against the schema:

```{r schema_validate1}
validate(
  schema,
  '{"productId": 1, "productName": "A green door", "price": 12.50}'
)
```

Note that instance can also be an R list directly (will be converted to JSON for validation):
```{r schema_validate2}
validate(
  schema,
  list(productId = 1, productName = "A green door", price = 12.50)
)
```

The JSON object is valid, as it adheres to the schema.

However, the following JSON object is not valid:

```{r schema_validate3}
validate(
  schema,
  '{"productId": "1", "productName": "A green door", "price": 0}'
)
res <- validate(
  schema,
  '{"productId": "1", "productName": "A green door", "price": 0}',
  style = "basic"
)
str(res)
```

